\documentclass[a4paper,12pt]{article}

\input{./preamble.tex}


\title{Ensemble Data Assimilation for Meshless Methods}
\author[1,2]{Marius Duvillard}
\author[1]{Loïc Giraldi}
\author[3]{Olivier Le Maître}
\affil[1]{CEA, DES, IRESNE, DEC, SESC, LMCP, Cadarache, F-13108 Saint-Paul-Lez-Durance, France}
\affil[3]{CNRS, Inria, Centre de Mathématiques Appliquées, Ecole Polytechnique, IPP, Route de Saclay, 91128, Palaiseau Cedex, France}
\affil[2]{Centre de Mathématiques Appliquées, Ecole Polytechnique, IPP, Route de Saclay, 91128, Palaiseau Cedex, France}
\date{}

% \usepackage[
%     backend=biber,
%     style=alphabetic,
%     sorting=ynt
% ]{biblatex}
% \usepackage{biblatex}
% \addbibresource{biblio.bib}
% \bibliographystyle{elsarticle-num}
\bibliographystyle{plain}
% \bibliographystyle{ACM-Reference-Format}


\begin{document}
% \begin{frontmatter}
\maketitle

\begin{abstract}
    This study presents a novel approach for integrating data assimilation techniques into meshless simulations using the Ensemble Kalman Filter (EnKF). Combining EnKF with meshless methods facilitates an efficient estimation across various problems utilizing a Lagrangian solution discretization. Two specific methodologies have been introduced to either complete the analysis on a new grid of particles or on the forecast particle discretization of each member. These methods were initially evaluated using a one-dimensional advection-diffusion model with periodic boundaries. Subsequently, they were applied to a more complex two-dimensional inviscid flow problem, solved via the Vortex-In-Cell (VIC) method. In the one-dimensional scenario, the performance of these filters was benchmarked against a grid-based assimilation filter. In the two-dimensional case, the study demonstrated the feasibility of applying these methods in more intricate scenarios.

\end{abstract}

{\bf Keywords:} Meshless Methods, Particle-based Method, Data Assimilation, EnKF, Ensemble Methods, Vortex Methods.
% \begin{keyword}
%     meshless methods \sep data assimilation \sep EnKF \sep Ensemble Methods 
% \end{keyword}

% \end{frontmatter}

\tableofcontents

\input{./introduction.tex}
\input{./bg-da.tex}
\input{./bg-part.tex}
\input{./methods.tex}
\input{./app-1d.tex}
\input{./app-2d.tex}
\input{./conclusion.tex}
\input{./appendix.tex}
% \section*{References}

\bibliography{biblio}

% \nocite{*}
% \printbibliography

\end{document}