\documentclass[a4paper,12pt]{article}

\input{./preamble.tex}


\title{Ensemble Data Assimilation for Meshless Methods}
\author[1,2]{Marius Duvillard}
\author[1]{Loïc Giraldi}
\author[3]{Olivier Le Maître}
\affil[1]{CEA, DES, IRESNE, DEC, SESC, LMCP, Cadarache, F-13108 Saint-Paul-Lez-Durance, France}
\affil[3]{CNRS, Inria, Centre de Mathématiques Appliquées, Ecole Polytechnique, IPP, Route de Saclay, 91128, Palaiseau Cedex, France}
\affil[2]{Centre de Mathématiques Appliquées, Ecole Polytechnique, IPP, Route de Saclay, 91128, Palaiseau Cedex, France}
\date{}

% \usepackage[
%     backend=biber,
%     style=alphabetic,
%     sorting=ynt
% ]{biblatex}
% \usepackage{biblatex}
% \addbibresource{biblio.bib}
% \bibliographystyle{elsarticle-num}
\bibliographystyle{plain}
% \bibliographystyle{ACM-Reference-Format}


\begin{document}
% \begin{frontmatter}
\maketitle

\begin{abstract}
    This study presents a novel approach for integrating data assimilation techniques into meshless simulations using the Ensemble Kalman Filter. If data assimilation methods have been apply on Eulerian simulations for long, there have never been properly used in the context of a Lagrangian solution discretization. Two specific methodologies are introduced to complete the analysis. The first one is based on the use of an intermediary Eulerian transformation combining a projection and a remeshing process. The second is a purely Lagrangian  scheme useful when remeshing is not adapted. These methods are evaluated using a one-dimensional advection-diffusion model with periodic boundaries. Subsequently, assimilation schere are applied to a non-linear two-dimensional inviscid flow problem, solved via the Vortex-In-Cell method. In the one-dimensional scenario, the performance of these filters is benchmarked against a grid-based assimilation filter. In the two-dimensional case, the study demonstrates the feasibility of applying these methods in more intricate scenarios.

\end{abstract}

{\bf Keywords:} Meshless Methods, Particle-based Method, Data Assimilation, EnKF, Ensemble Methods, Vortex Methods.
% \begin{keyword}
%     meshless methods \sep data assimilation \sep EnKF \sep Ensemble Methods 
% \end{keyword}

% \end{frontmatter}

\tableofcontents
\input{./introduction.tex}
\input{./bg-da.tex}
\input{./bg-part.tex}
\input{./methods.tex}
\input{./app-1d.tex}
\input{./app-2d.tex}
% \input{./conclusion.tex}
\input{./appendix.tex}
% \section*{References}

\bibliography{C:/Users/md266594/mar-dvd-thesis/articles/new_part_enkf/biblio}

% \nocite{*}
% \printbibliography

\end{document}